---
title: "Week3Day2 Factors"
author: "Dave and Mel"
date: "May 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## Factor Analysis

find corrplot snippets

Decide on approaches and tools


```{r cars}
library("corrplot")
library("ggplot2")
library("psych")
library("GPArotation")
library("Rmisc")
library("dplyr")
```

## PCA

use rnorm for 3 data frames
play with random data frames generated from rnorm

write function noisyindicators
 - note variable(i) = a1F1 + a2F2 + .... + akFk + error(i)
 - resolve confusion

make dataframe noisies

run PCA

Plot PCA versus noisies and compare to pca vs. factors

```{r PCA}

set.seed(1)
factors = data.frame(matrix(rnorm(300) , ncol = 3, nrow = 100))
names(factors) = c('X','Y','Z')

# y = a*x + error
# 

noisyIndicators = function(feature, name, k, correlation){
  
  result = vector(mode="list", k)
  for(i in 1:k) {
    error = rnorm(length(feature))
    proxy = correlation * feature + error
    result[[i]] = proxy
  }
  names(result) = paste(name, 1:k, sep="_")
  return(as.data.frame(result))
}

# 4 noisy proxies to X
# 3 noisy proxies to Y
noisies = cbind(noisyIndicators(factors$X, "X", 4, .9),
               noisyIndicators(factors$Y, "Y", 3, -.5))

# plot out the correlation between the true vars and proxies 
corrplot( cor(factors, noisies), cl.pos="n")

pca = prcomp(factors)
plot(pca$sdev)

# the pcas are still correlated with noisy Xs and noisy Ys,
# but less so, or in the opposite direction if noisies corr is negative
corrplot(pca$rotation)
corrplot(cor(pca$x,noisies))

noisy_pca = prcomp(noisies)

corrplot(noisy_pca$rotation, title = "Noisy PCA rotations")

#PC 1 of factors disappeared in the noisies... because it was correlated with Z
#PC 2 and 3 still there. PC2 is weaker, it correlate with Y which has a weakercorrelation with noisies

corrplot(cor(noisy_pca$x, pca$x), title="Noisy PCA vs. Factor PCA")

faa = fa(noisies, nfactors = 2, rotate="varimax")

# factor one correspond to Xs, factor two to Ys but more noisily
corrplot(faa$loadings)

# correlation between guessed factors and real factors
corrplot(cor(faa$scores, factors))

```

## Factor Analysis

Look at Drunkard's walk

get correlation on modeled factors and true factors

50 variables:
  X*runif(1) + Y*runif(2) + Z*runif(3) + 0.5*error
    - error normally distributed with SD 1

PCA on generated variables

plot 3 principal compenents vs XYZ correlation

factor analysis with k= 3 and rotate = 'varimax'

corr(modeled factors, true factors

```{r}
library(psych)

df = data.frame()

#initialize a holder for the 50 variables
varholder = vector('list',50L)

#populate
for (col in 1:50) {
  var = (factors$X * runif(1) + 
         factors$Y * runif(1) + 
         factors$Y * runif(1) + 
         factors$Z * runif(1) + 
         factors$Z * runif(1) + 
         factors$Z * runif(1) + 
         0.5 * rnorm(100))
  varholder[[col]] = var
}

df = as.data.frame(varholder)
names(df) = paste('var', 1:50, sep="")

pcafifty = prcomp(df, scale. = TRUE)
plot(pcafifty$sdev, xlab = 'a', ylab = 'b')

# rotations
corrplot(pcafifty$rotation[,1:3], is.corr=FALSE)



```

Factor Z is the first principal component. It's against 3 random numers, so it's an axis
with lots of variation available based on it.

```{r}
corrplot(cor(pcafifty$x[,1:4], factors))


fafifty = fa(df, nfactors = 3, rotate="varimax")

corrplot(cor(fafifty$scores, factors))



```
## Oblique factor analysis

W = 0.5*X + Y

corr(W, Y)

noisyIndicators(10 indicators from X)
  noisy(4 with W)
  correlations = .8 in both
  
corr(noisyW, noisyX) ? not clear on instructions

fa(nfactors = 2, rotate = "varimax")
fa(nfac=2, rotate = "oblimin")

? construct tall dataframe to plot?

how would you compare the correlations?
results with noisy indicators
results with true factors
either of those dependent on style of rotation

```{r oblique}

W = 0.5*factors$X + factors$Y

cor(W,factors$Y)

noiOblique = cbind(noisyIndicators(factors$X, "X", 10, .8),
               noisyIndicators(W, "W", 4, .8))

xandw = matrix(c(W,factors$X), ncol = 2)

corrplot(cor(noiOblique,xandw))

# test out oblique factor analysis

varmax = fa(noiOblique, nfactors = 2, rotate="varimax")
oblim = fa(noiOblique, nfactors = 2, rotate="oblimin")

colnames(oblim$loadings) = c("oblim_MR1", "oblim_MR2")
vandoblim = cbind(varmax$loadings, oblim$loadings)

corrplot(vandoblim, is.corr=FALSE, cl.pos = "n" )

```

### Speed dating data

clean up speed data frame

```{r speed dating}
speed_df = read.csv("speeddating-aggregated.csv")

features = select(speed_df, sports:yoga)

```
loop over 1,2,3,4 nfactor
  Do factor analysis on activities w/varimax
  do with oblimin

look at what those results are like, figure out how to get on
  one or two corrplots

corrplot factor loadings


```{r plot factor loadings, eval=FALSE, include=FALSE}
cache = list(varimax = vector("list", 4),
             oblimin = vector("list", 4))

for (i in 1:4) {
  cache$varimax[[i]] = fa(features, nfactors=i, rotate="varimax")
  cache$oblimin[[i]] = fa(features, nfactors=i, rotate="oblimin")
}


vdf = as.data.frame(matrix(0, nrow=ncol(features), ncol=sum(1:4)))
odf = as.data.frame(matrix(0, nrow=ncol(features), ncol=sum(1:4)))
coln = 1
for (i in 1:4) {
  for (j in 1:i) {
    vdf[[coln]] = cache$varimax[[i]]$loadings[,j]
    odf[[coln]] = cache$oblimin[[i]]$loadings[,j]
    colnames(vdf)[coln] = paste("vrmax",i,"MR",j,sep="")
    colnames(odf)[coln] = paste("oblim",i,"MR",j,sep="")
    coln = coln + 1
  }
}
rownames(vdf) = colnames(features)
rownames(odf) = colnames(features)
corrplot(as.matrix(vdf), is.corr=FALSE, cl.pos="n")
corrplot(as.matrix(odf), is.corr=FALSE, cl.pos="n")


```
#### Varimax

*First component* with vdf is invariable about museums, art, and theatre.
*Second component* is kind of tv, sports, gaming the first run it goes on, but decoheres after.

#### Oblimin

*First component* is the same
*Second component* is kinda spread thing when it first comes up, gets a bit more concentrated on the third around tv through shoppping, but then on the fourth becomes clearly about concerts and music.
*Third component* is clearly sports/tv/exercise

```{r continued oblique factors}
corrplot(cor(vdf, odf))

# the two types of component analysis correspond really well to their counterparts, up till 3 and 4 which look switched.


```
### Big Five personality data

downlaod big5 dataset

explore
clean 

restrict to 50 questions?
get gender as target

PCA on 50 questions
factor analysis n=5 on 50 questions

what is compare results?
interpret both loadings?

corrplot features against rotations
corrplot features against loadngs
corrplot rotations against loadings

glm(family=binomal) predict gender on 5 factors
glm(family=binomal) predict gender on 50 qs
display coefficients (corr first, filter then corr second)
? discuss differences, write notes

ask about 10 questions for each trait I don't get this
mark which question to each trait with dummies
train on trait# factors, look at corr between $x thing and actual traits

 OH! split into trait# of datasets, train 1 factor on each
 look up subfactors of each trait
 factoranalysis those numbers and look at loadings to see if questions make sense